Transactional Memory Concurrency Verification with Landslide

### 2018-03-15 htm/paper.tex
minor reorganization of bunch of stuff

\begin{table}[t]
	\begin{center}
	\begin{tabular}{l|c|c|c}
		load & cputime (s) & vs self avg & vs baseline \\
		\hline
		{\sf none	} & 14.95 $\pm$ 0.17	& 0.99-1.02x	& (baseline)	\\
		{\sf vid-L	} & 15.13 $\pm$ 0.10	& 0.99-1.01x	& 1.01x	\\
		{\sf ff/c	} & 15.55 $\pm$ 0.16	& 0.99-1.02x	& 1.04x	\\
		{\sf sm5	} & 16.63 $\pm$ 0.07	& 0.99-1.01x	& 1.11x \\
		{\sf ff/c+vid-S	} & 17.09 $\pm$ 0.34	& 0.98-1.05x	& 1.14x	\\
		{\sf ls		} & 19.11 $\pm$ 0.32	& 0.97-1.03x	& 1.28x	\\
		% simply love title screen
		%{\sf sm5+sl}	& 19.18 $\pm$ 0.32	& 0.97-1.02x	& 1.28x \\
		{\sf ff/c+ls	} & 19.66 $\pm$ 0.50	& 0.96-1.02x	& 1.31x	\\
	\end{tabular}
	\end{center}
	\caption{Performance variance on {\tt counter}(2,2) with other programs running on the test machine.
		{\sf vid-L} is full-screen video (played locally with {\tt mplayer}),
		{\sf ff/c} is Firefox and Chrome (idle, $\le$20 tabs),
		{\sf sm5} is StepMania 5.1 (during gameplay \cite{itg2}),
		{\sf vid-S} is full-screen video (streamed via Crunchyroll),
		{\sf ls} is a 2nd instance of Landslide.
		Average of 10 samples, $\pm N$ is 1 stddev.
	}
	\label{tab:variants}
\end{table}

@@

\begin{table}[t]
	\begin{center}
	\begin{tabular}{l|r|r}
		Landslide mode & total int's & cputime (s) \\
		\hline
		{\sf verif} ({\tt -M})		& 403 & 14.95 $\pm$ 00.17	\\
		{\sf reinstrument}		& 403 & 24.72 $\pm$ 00.10	\\
		{\sf quicksand} (no {\tt -M})	& 2128 & 120.53 $\pm$ 18.62	\\
	\end{tabular}
	\end{center}
	\caption{Performance variance on {\tt counter}(2,2) in various modes.
		Average of 10 samples, $\pm N$ is 1 stddev.
	}
	\label{tab:variants2}
\end{table}

We conclude that MC performance evaluations must address
experimental environment variables
%at least enough
to ensure consistent performance between runs.
Considering recent news on cross-VM cache attacks \cite{meltdown,spectre},
we also conjecture that cloud-based virtual machines cannot be trusted for low-variance performance measurements.
So doing, multiple samples and error bars are then necessary
only when using nondeterministic search ordering strategies
such as Iterative Deepening.
Otherwise, considering the low variance shown in Table~\ref{tab:variants},
they need be shown only for a token small test
to promise the reader the same consistency in the larger tests
(especially if the time tradeoff to measure them would sacrifice testing larger state spaces to begin with).
% We also advise the busy researcher to torrent their anime rather than stream it. % lol can't put this in a paper

Outside of Table~\ref{tab:variants}, we measured all performance numbers
with Firefox and Chrome as the only other significant machine load.
(we gotta stay amused somehow).
The numbers of interleavings in each state space are, of course, deterministic and do not vary across runs.

@@

Many such branches may be equivalent;
for example,
explicit and conflict aborts need not be tested separately in transactions
whose failure paths do not distinguish the cause,
and the perennial {\tt \_XABORT\_RETRY} abort can be skipped entirely
if the client abstracts it away behind a retry loop.
In fact, applying both reductions simultaneously
produces a state space smaller than the original;
testing them by hand (trusting their soundness by visual inspection)
on {\tt map\_basic}(2,1) reduced its 28635 interleavings to 11577.
%We tested both reductions by hand, trusting their soundness by visual inspection,
%on {\tt map\_basic}(2,1),
%and preview the resulting state space sizes in Table~\ref{tab:abort-codes}.
%
%\begin{table}[h]
%	\caption{...}
%	\label{tab:abort-codes}
%\end{table}
%
Future work could use static or dynamic flow analysis to
identify such reduction opportunities automatically;
for now, this feature is disabled by default but accessible via the {\tt -A} command-line option
to Landslide (in addition to {\tt -X}).

{\bf Relaxed memory orderings.}
Section~\ref{sec:design}'s formalization of thread interleavings does not account for read/write reorderings
possible on relaxed memory consistency architectures \cite{memory-consistency-models}.
In fact,
even after \cite{htm-mario}'s proposed fix,
our running example program is still incorrect on Total Store Ordering (TSO) architectures such as x86.
Despite stores being totally-ordered, x86 may still reorder stores after subsequent loads.
Accordingly, an execution of $8,9a,9b,9c$ in Figure~\ref{fig:htm-fixed}
may be locally visible to another thread as $9a,8,9b,9c$,
and hence an apparent interleaving of
\[
	\tiat{1},\tjat{1-5},\tiat{7},\underline{\tiat{9a}},\tkat{1-5},\underline{\tiat{8}},\tiat{9b-B}
\]
is possible
(reordered accesses underlined for emphasis).
An {\tt mfence} barrier is needed between lines 8 and 9 to solve this problem on TSO \cite{tsx-need-barrier}.
On Partial Store Ordering (PSO) architectures, even more barriers may be necessary.

>>>>

\begin{table}[h]
	\begin{center}
	\begin{tabular}{l|c|c|c}
		load & cputime (s) & vs self avg & vs baseline \\
		\hline
		{\sf none	} & 14.95 $\pm$ 0.17	& 0.99-1.02x	& (baseline)	\\
		{\sf vid-L	} & 15.13 $\pm$ 0.10	& 0.99-1.01x	& 1.01x	\\
		{\sf ff/c	} & 15.55 $\pm$ 0.16	& 0.99-1.02x	& 1.04x	\\
		{\sf sm5	} & 16.63 $\pm$ 0.07	& 0.99-1.01x	& 1.11x \\
		{\sf ff/c+vid-S	} & 17.09 $\pm$ 0.34	& 0.98-1.05x	& 1.14x	\\
		{\sf ls		} & 19.11 $\pm$ 0.32	& 0.97-1.03x	& 1.28x	\\
		% simply love title screen
		%{\sf sm5+sl}	& 19.18 $\pm$ 0.32	& 0.97-1.02x	& 1.28x \\
		{\sf ff/c+ls	} & 19.66 $\pm$ 0.50	& 0.96-1.02x	& 1.31x	\\
	\end{tabular}
	\end{center}
	\caption{Performance variance on {\tt counter}(2,2) with other programs running on the test machine.
		{\sf vid-L} is full-screen video (played locally with {\tt mplayer}),
		{\sf ff/c} is Firefox and Chrome (idle, $\le$20 tabs),
		{\sf sm5} is StepMania 5.1 (during gameplay \cite{itg2}),
		{\sf vid-S} is full-screen video (streamed via Crunchyroll),
		{\sf ls} is a 2nd instance of Landslide.
		Average of 10 samples, $\pm N$ is 1 stddev.
	}
	\label{tab:variants}
\end{table}

@@

\begin{table}[h]
	\begin{center}
	\begin{tabular}{l|r|r}
		Landslide mode & total int's & cputime (s) \\
		\hline
		{\sf verif} ({\tt -M})		& 403 & 14.95 $\pm$ 00.17	\\
		{\sf reinstrument}		& 403 & 24.72 $\pm$ 00.10	\\
		{\sf quicksand} (no {\tt -M})	& 2128 & 120.53 $\pm$ 18.62	\\
	\end{tabular}
	\end{center}
	\caption{Performance variance on {\tt counter}(2,2) in various modes.
		Average of 10 samples, $\pm N$ is 1 stddev.
	}
	\label{tab:variants2}
\end{table}

We conclude that MC performance evaluations must address
experimental environment variables
%at least enough
to ensure consistent performance between runs.
So doing, multiple samples and error bars are then necessary
only when using nondeterministic search ordering strategies
such as Iterative Deepening.
Otherwise, considering the low variance shown in Table~\ref{tab:variants},
they need be shown only for a token small test
to promise the reader the same consistency in the larger tests
(especially if the time tradeoff to measure them would sacrifice testing larger state spaces to begin with).
% We also advise the busy researcher to torrent their anime rather than stream it. % lol can't put this in a paper
%% i'm not actually sure this fits..
%Finally, considering recent news on cross-VM cache attacks \cite{meltdown,spectre},
%we also conjecture that cloud-based virtual machines cannot be trusted for low-variance performance measurements.

Outside of Table~\ref{tab:variants}, we fixed our environment by measuring all performance numbers
with Firefox and Chrome as the only other significant machine load.
%(we gotta stay amused somehow).
We believe the exponential differences among completion times justifies the absence of error bars,
which would be expected to show 2\% variance if extrapolating from this experiment.
The numbers of interleavings in each state space are, of course, deterministic and do not vary across runs.

@@

Many such branches may be equivalent;
for example,
explicit and conflict aborts need not be tested separately in transactions
whose failure paths do not distinguish the cause,
and the perennial {\tt \_XABORT\_RETRY} abort can be skipped entirely
if the client abstracts it away behind a retry loop.
In fact, applying both reductions simultaneously
produces a state space smaller than the original;
testing them by hand (trusting their soundness by visual inspection)
on {\tt map\_basicer}(2,1) reduced its 28635 interleavings to 11577
(although {\tt map\_basic}(2,1) lies still beyond reach).
%We tested both reductions by hand, trusting their soundness by visual inspection,
%on {\tt map\_basic}(2,1),
%and preview the resulting state space sizes in Table~\ref{tab:abort-codes}.
%
%\begin{table}[h]
%	\caption{...}
%	\label{tab:abort-codes}
%\end{table}
%
Future work could use static or dynamic flow analysis to
identify such reduction opportunities automatically;
for now, this feature is disabled by default but accessible via the {\tt -A} command-line option
to Landslide (in addition to {\tt -X}).

{\bf Relaxed memory orderings.}
Section~\ref{sec:design}'s formalization of thread interleavings does not account for read/write reorderings
possible on relaxed consistency architectures \cite{memory-consistency-models}.
In fact,
even after \cite{htm-mario}'s proposed fix,
our running example program is still incorrect on Total Store Ordering (TSO) architectures such as x86.
Despite stores being totally-ordered, x86 may still reorder stores after subsequent loads.
Accordingly, an execution of $8,9a,9b,9c$ in Figure~\ref{fig:htm-fixed}
may be locally visible to another thread as $9a,8,9b,9c$,
and hence an apparent interleaving of
\[
	\tiat{1},\tjat{1-5},\tiat{7},\underline{\tiat{9a}},\tkat{1-5},\underline{\tiat{8}},\tiat{9b-B}
\]
is possible
(reordered accesses underlined for emphasis).
An {\tt mfence} barrier is needed between lines 8 and 9 to solve this problem on TSO \cite{tsx-need-barrier}.
On Partial Store Ordering (PSO) architectures, even more barriers may be necessary.
